---
title: 'Homework 4: Machine Learning'
author: "Ignacio Landerreche"
date: '`r Sys.Date()`'
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r}
#| label: load-libraries
#| echo: false # This option disables the printing of code (only output is displayed).
#| message: false
#| warning: false
options(scipen = 999) #disable scientific notation
library(tidyverse)
library(tidymodels)
library(GGally)
library(sf)
library(leaflet)
library(janitor)
library(rpart.plot)
library(here)
library(scales)
library(vip)
```


# The Bechdel Test

https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/

The [Bechdel test](https://bechdeltest.com) is a way to assess how women are depicted in Hollywood movies.  In order for a movie to pass the test:

1. It has to have at least two [named] women in it
2. Who talk to each other
3. About something besides a man

There is a nice article and analysis you can find here https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/
We have a sample of 1394 movies and we want to fit a model to predict whether a film passes the test or not.

```{r read_data}

bechdel <- read_csv(here::here("data", "bechdel.csv")) %>% 
  mutate(test = factor(test)) 
glimpse(bechdel)

```
How many films fail/pass the test, both as a number and as a %?
_ANSWER(%):_ Fail (55%),Pass (45%)

```{r}
bechdel %>% 
  count(test) %>% 
  mutate(freq_perc=round(n*100/sum(n),2))

```


## Movie scores
```{r}
ggplot(data = bechdel, aes(
  x = metascore,
  y = imdb_rating,
  colour = test
)) +
  geom_point(alpha = .3, size = 3) +
  scale_colour_manual(values = c("tomato", "olivedrab")) +
  labs(
    x = "Metacritic score",
    y = "IMDB rating",
    colour = "Bechdel test"
  ) +
 theme_light()
```


# Split the data
```{r}
# **Split the data**

set.seed(123)

data_split <- initial_split(bechdel, # updated data
                           prop = 0.8, 
                           strata = test)

bechdel_train <- training(data_split) 
bechdel_test <- testing(data_split)
```

Check the counts and % (proportions) of the `test` variable in each set.
```{r}
#Number of rows for training and percentage of each test
nrow(bechdel_train)

bechdel_train %>% 
  count(test) %>% 
  mutate(freq_perc=round(n*100/sum(n),2))

#Number of rows for testing and percentage of each test
nrow(bechdel_test)

bechdel_test %>% 
  count(test) %>% 
  mutate(freq_perc=round(n*100/sum(n),2))
```

## Feature exploration

## Any outliers? 
Many, specially in intgross and domgross. Particularily there is a single movie which is very big outlier in the "Fail" in domgross_2013

```{r}

bechdel %>% 
  select(test, budget_2013, domgross_2013, intgross_2013, imdb_rating, metascore) %>% 

    pivot_longer(cols = 2:6,
               names_to = "feature",
               values_to = "value") %>% 
  ggplot()+
  aes(x=test, y = value, fill = test)+
  coord_flip()+
  geom_boxplot()+
  facet_wrap(~feature, scales = "free")+
  theme_bw()+
  theme(legend.position = "none")+
  labs(x=NULL,y = NULL)

```

## Scatterplot - Correlation Matrix
Some notes:
1. Regarding the "test" variable:
  a) there is a left skew in both pass and fail for the budget which suggest that the sample consists of movies with high budget.
  b) There seems to be a more symmetrical distribution with the scores for both pass and fail test
2. Budget variable:
  a)  It seems to be that there is a very high outlier for a fail film, and in frequency there seems to be more outliers in the pass films than the fail films.
  b) distribution is right skewed with budget
  c) no clear relationship with scores
3. Domgross and intgross: looks like they have a very high linear correlation (0.93), and both are highly         correlated with budget too
4. IMDB_Rating and metascore have high correlation too (0.737)
  

```{r, warning=FALSE, message=FALSE}
bechdel %>% 
  select(test, budget_2013, domgross_2013, intgross_2013, imdb_rating, metascore)%>% 
  ggpairs(aes(colour=test), alpha=0.2)+
  theme_bw()
```


## Categorical variables
One can see the following conclusions by genre:
1. There are some genres that failed 100% of the times, however sample size in both is quite small compared to other genres (Documentary and Sci-FI).
2. From the rest of the genres, the one that seems to fail the most with a sample size >100 is Action with 70% of the times which seems intuitive.
3. The genre that passed the most with a sample size >100 is Comedy with a failure rate of 42%

One can see the following conclusions by rating:
1.Not many movies are classified from G and NC-17, which might then not be a good estimation for the failing test.
2. For the rest of the classifications, the proportions seem to be stable among failure and passing rate. 

```{r}
bechdel %>% 
  group_by(genre, test) %>%
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n))

#Filtering only the failing and ordering them
bechdel %>% 
  group_by(genre, test) %>%
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n)) %>% 
  filter(test=="Fail") %>% 
  arrange(desc(prop))
  
 
bechdel %>% 
  group_by(rated, test) %>%
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n))
```

# Train first models. `test ~ metascore + imdb_rating`

```{r}
lr_mod <- logistic_reg() %>% 
  set_engine(engine = "glm") %>% 
  set_mode("classification")

lr_mod


tree_mod <- decision_tree() %>% 
  set_engine(engine = "C5.0") %>% 
  set_mode("classification")

tree_mod 
```

```{r}


lr_fit <- lr_mod %>% # parsnip model
  fit(test ~ metascore + imdb_rating, # a formula
    data = bechdel_train # dataframe
  )

tree_fit <- tree_mod %>% # parsnip model
  fit(test ~ metascore + imdb_rating, # a formula
    data = bechdel_train # dataframe
  )
```

## Logistic regression

```{r}
lr_fit %>%
  broom::tidy()

lr_preds <- lr_fit %>%
  augment(new_data = bechdel_train) %>%
  mutate(.pred_match = if_else(test == .pred_class, 1, 0))

```

### Confusion matrix

```{r}
lr_preds %>% 
  conf_mat(truth = test, estimate = .pred_class) %>% 
  autoplot(type = "heatmap")


```


## Decision Tree
```{r}
tree_preds <- tree_fit %>%
  augment(new_data = bechdel) %>%
  mutate(.pred_match = if_else(test == .pred_class, 1, 0)) 


```

```{r}
tree_preds %>% 
  conf_mat(truth = test, estimate = .pred_class) %>% 
  autoplot(type = "heatmap")
```

## Draw the decision tree

```{r}
draw_tree <- 
    rpart::rpart(
        test ~ metascore + imdb_rating,
        data = bechdel_train, # uses data that contains both birth weight and `low`
        control = rpart::rpart.control(maxdepth = 5, cp = 0, minsplit = 10)
    ) %>% 
    partykit::as.party()
plot(draw_tree)

```

# Cross Validation

Run the code below. What does it return? 
_ANSWER_: The Vfold is a cross-validation feature from R, it divides the data in 10 different bins with equal-sized samples. Each of the 10 partitions will be chosen to be the testing set once with the other 9 partitions used for training. 

```{r}
set.seed(123)
bechdel_folds <- vfold_cv(data = bechdel_train, 
                          v = 10, 
                          strata = test)
bechdel_folds
```

## `fit_resamples()`

Trains and tests a resampled model.

```{r}
lr_fit <- lr_mod %>%
  fit_resamples(
    test ~ metascore + imdb_rating,
    resamples = bechdel_folds
  )


tree_fit <- tree_mod %>%
  fit_resamples(
    test ~ metascore + imdb_rating,
    resamples = bechdel_folds
  )
```


## `collect_metrics()`

Unnest the metrics column from a tidymodels `fit_resamples()`
```{r}

collect_metrics(lr_fit)
collect_metrics(tree_fit)
```


```{r}
tree_preds <- tree_mod %>% 
  fit_resamples(
    test ~ metascore + imdb_rating, 
    resamples = bechdel_folds,
    control = control_resamples(save_pred = TRUE) #<<
  )

# What does the data for ROC look like? Very bad, it is almost like the line meaning it is almost random.
tree_preds %>% 
  collect_predictions() %>% 
  roc_curve(truth = test, .pred_Fail)  

# Draw the ROC
tree_preds %>% 
  collect_predictions() %>% 
  roc_curve(truth = test, .pred_Fail) %>% 
  autoplot()

```


# Build a better training set with `recipes`

## Preprocessing options

- Encode categorical predictors
- Center and scale variables
- Handle class imbalance
- Impute missing data
- Perform dimensionality reduction 
- ... ...

## To build a recipe

1. Start the `recipe()`
1. Define the variables involved
1. Describe **prep**rocessing [step-by-step]

## Collapse Some Categorical Levels

Do we have any `genre` with few observations?  Assign genres that have less than 3% to a new category 'Other'


```{r}
#| echo = FALSE
bechdel %>% 
  count(genre) %>% 
  mutate(genre = fct_reorder(genre, n)) %>% 
  ggplot(aes(x = genre, 
             y = n)) +
  geom_col(alpha = .8) +
  coord_flip() +
  labs(x = NULL) +
  geom_hline(yintercept = (nrow(bechdel_train)*.03), lty = 3)+
  theme_light()
```


```{r}
movie_rec <-
  recipe(test ~ .,
         data = bechdel_train) %>%
  
  # Genres with less than 5% will be in a catewgory 'Other'
    step_other(genre, threshold = .03) 
```
  

## Before recipe

```{r}
#| echo = FALSE
bechdel_train %>% 
  count(genre, sort = TRUE)
```


## After recipe

```{r}
movie_rec %>% 
  prep() %>% 
  bake(new_data = bechdel_train) %>% 
  count(genre, sort = TRUE)
```

## `step_dummy()`

Converts nominal data into numeric dummy variables

```{r}
#| results = "hide"
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_dummy(all_nominal_predictors()) 

movie_rec 
```

## Let's think about the modelling 

What if there were no films with `rated` NC-17 in the training data?

-  Will the model have a coefficient for `rated` NC-17?
-  What will happen if the test data includes a film with `rated` NC-17?

## `step_novel()`

Adds a catch-all level to a factor for any new values not encountered in model training, which lets R intelligently predict new levels in the test set.

```{r}

movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal_predictors) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal_predictors()) 

```


## `step_zv()`

Intelligently handles zero variance variables (variables that contain only a single value)

```{r}
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal(), -all_outcomes()) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_numeric(), -all_outcomes()) 
  
```


## `step_normalize()`

Centers then scales numeric variable (mean = 0, sd = 1)

```{r}
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal(), -all_outcomes()) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_numeric(), -all_outcomes())  %>% 
  step_normalize(all_numeric()) 

```


## `step_corr()`

Removes highly correlated variables

```{r}
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal(), -all_outcomes()) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_numeric(), -all_outcomes())  %>% 
  step_normalize(all_numeric()) 

movie_rec
```


# Define different models to fit

```{r}
## Model Building

# 1. Pick a `model type`
# 2. set the `engine`
# 3. Set the `mode`: regression or classification

# Logistic regression
log_spec <-  logistic_reg() %>%  # model type
  set_engine(engine = "glm") %>%  # model engine
  set_mode("classification") # model mode

# Show your model specification
log_spec

# Decision Tree
tree_spec <- decision_tree() %>%
  set_engine(engine = "C5.0") %>%
  set_mode("classification")

tree_spec

# Random Forest
library(ranger)

rf_spec <- 
  rand_forest() %>% 
  set_engine("ranger", importance = "impurity") %>% 
  set_mode("classification")


# Boosted tree (XGBoost)
library(xgboost)

xgb_spec <- 
  boost_tree() %>% 
  set_engine("xgboost") %>% 
  set_mode("classification") 

# K-nearest neighbour (k-NN)
knn_spec <- 
  nearest_neighbor(neighbors = 4) %>% # we can adjust the number of neighbors 
  set_engine("kknn") %>% 
  set_mode("classification") 
```


# Bundle recipe and model with `workflows`


```{r}
log_wflow <- # new workflow object
 workflow() %>% # use workflow function
 add_recipe(movie_rec) %>%   # use the new recipe
 add_model(log_spec)   # add your model spec

# show object
log_wflow


## A few more workflows

tree_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(tree_spec) 

rf_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(rf_spec) 

xgb_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(xgb_spec)

knn_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(knn_spec)

```

HEADS UP

1. How many models have you specified? _ANSWER_: 5 models
2. What's the difference between a model specification and a workflow? _ANSWER_ The workflows involves many steps (like choosing the model, adding a recipe, etc), in which the model specification is one step. However for other machine learning types one may need to have more steps within the workflow. 
3. Do you need to add a formula (e.g., `test ~ .`)  if you have a recipe? _ANSWER_ No, because the recipe already has the formula embedded.


# Model Comparison

You now have all your models. Adapt the code from slides `code-from-slides-CA-housing.R`, line 400 onwards to assess which model gives you the best classification. 


```{r}
## Logistic regression results
log_res <- log_wflow %>% 
  fit_resamples(
    resamples = bechdel_folds, 
    metrics = metric_set(
      recall, precision, f_meas, accuracy,
      kap, roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)) 

# Show average performance over all folds (note that we use log_res):
log_res %>%  collect_metrics(summarize = TRUE)

# Show performance for every single fold:
log_res %>%  collect_metrics(summarize = FALSE)



## `collect_predictions()` and get confusion matrix{.smaller}

log_pred <- log_res %>% collect_predictions()

log_pred %>%  conf_mat(test, .pred_class) 

log_pred %>% 
  conf_mat(test, .pred_class) %>% 
  autoplot(type = "mosaic") +
  geom_label(aes(
      x = (xmax + xmin) / 2, 
      y = (ymax + ymin) / 2, 
      label = c("TP", "FN", "FP", "TN")))


log_pred %>% 
  conf_mat(test, .pred_class) %>% 
  autoplot(type = "heatmap")


## ROC Curve

log_pred %>% 
  group_by(id) %>% # id contains our folds
  roc_curve(test, .pred_Fail) %>% 
  autoplot()

```

```{r}
## Decision Tree results

tree_res <-
  tree_wflow %>% 
  fit_resamples(
    resamples = bechdel_folds,
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 

tree_res %>%  collect_metrics(summarize = TRUE)
```

```{r}

## Random Forest

rf_res <-
  rf_wflow %>% 
  fit_resamples(
    resamples = bechdel_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 

rf_res %>%  collect_metrics(summarize = TRUE)
```

```{r}
## Boosted tree - XGBoost

xgb_res <- 
  xgb_wflow %>% 
  fit_resamples(
    resamples = bechdel_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 

xgb_res %>% collect_metrics(summarize = TRUE)

```


```{r}
## K-nearest neighbour
library(kknn)
knn_res <- 
  knn_wflow %>% 
  fit_resamples(
    resamples = bechdel_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 

knn_res %>% collect_metrics(summarize = TRUE)

```


```{r}
## Model Comparison

log_metrics <- 
  log_res %>% 
  collect_metrics(summarise = TRUE) %>%
  # add the name of the model to every row
  mutate(model = "Logistic Regression") 

tree_metrics <- 
  tree_res %>% 
  collect_metrics(summarise = TRUE) %>%
  mutate(model = "Decision Tree")

rf_metrics <- 
  rf_res %>% 
  collect_metrics(summarise = TRUE) %>%
  mutate(model = "Random Forest")

xgb_metrics <- 
  xgb_res %>% 
  collect_metrics(summarise = TRUE) %>%
  mutate(model = "XGBoost")

knn_metrics <- 
  knn_res %>% 
  collect_metrics(summarise = TRUE) %>%
  mutate(model = "Knn")

# create dataframe with all models
model_compare <- bind_rows(log_metrics,
                           tree_metrics,
                           rf_metrics,
                           xgb_metrics,
                           knn_metrics) 

#Pivot wider to create barplot
  model_comp <- model_compare %>% 
  select(model, .metric, mean, std_err) %>% 
  pivot_wider(names_from = .metric, values_from = c(mean, std_err)) 

# show mean are under the curve (ROC-AUC) for every model
model_comp %>% 
  arrange(mean_roc_auc) %>% 
  mutate(model = fct_reorder(model, mean_roc_auc)) %>% # order results
  ggplot(aes(model, mean_roc_auc, fill=model)) +
  geom_col() +
  coord_flip() +
  scale_fill_brewer(palette = "Blues") +
   geom_text(
     size = 3,
     aes(label = round(mean_roc_auc, 2), 
         y = mean_roc_auc + 0.08),
     vjust = 1
  )+
  theme_light()+
  theme(legend.position = "none")+
  labs(y = NULL)

```



```{r}
#I will choose the Random Forest model as it is the one with highest Mean ROC AUS

last_fit_rf <- last_fit(rf_wflow, 
                        split = data_split,
                        metrics = metric_set(
                          accuracy, f_meas, kap, precision,
                          recall, roc_auc, sens, spec))

last_fit_rf %>% collect_metrics(summarize = TRUE)

#Compare to training
rf_res %>% collect_metrics(summarize = TRUE)
```

```{r}

## Variable importance using `{vip}` package

library(vip)

last_fit_rf %>% 
  pluck(".workflow", 1) %>%
  extract_fit_parsnip() %>% 
  vip(num_features = 10) +
  theme_light()


## Final Confusion Matrix

last_fit_rf %>%
  collect_predictions() %>% 
  conf_mat(test, .pred_class) %>% 
  autoplot(type = "heatmap")


## Final ROC curve
last_fit_rf %>% 
  collect_predictions() %>% 
  roc_curve(test, .pred_Fail) %>% 
  autoplot()


```


# Details

-   Who did you collaborate with: Myself
-   Approximately how much time did you spend on this problem set: One hour
-   What, if anything, gave you the most trouble: Understanding the differences between workflow, models and fitting



